\chapter{Python Semantic Module}

\section{Overview}

The Python semantic module uses the \emph{Jython} package
available from \emph{http://jython.org/} to allow
Python expressions and scripts to be used in process definitions.
The Python tag library is imported using an XML namespace
declaration with the URI of \verb|werkflow:python|.

\begin{codelisting}
<process \hired{xmlns:python="werkflow:python"} \dots>
    \dots
\end{codelisting}

\section{\tag{python:selector}}

The \tag{python:selector} tag allows selection of messages based
upon their true Java class and an optional filtering expression.
This selector only works when the WFMS is configured to use a 
\verb|SimpleMessagingManager|.  

Each message object received by the messaging manager will subject 
to an \verb|instanceof| check against the class specified by 
the \verb|type| attribute.  If an optional filtering expression
is provided by the \verb|filter| attribute, it is evaluated
with the message bound to the \verb|message| variable.

\begin{codelisting}
<message-type \dots>
    \hired{<python:selector type="com.myco.messages.PoMessage"
                   filter="message.getType() == 'reject'"/>}
</message-type>
\end{codelisting}

\begin{attrDefs}
type		&	required	&			& Java message class name. \\
filter		&	optional	&			& Filtering Python expression. \\
\end{attrDefs}

\section{\tag{python:correlator}}

The \tag{python:correlator} tag takes a single \verb|test| attribute
that should contain a boolean Python expression determining if
the message matches the process case.  The containing message-reception
action may specify that the message is bound to a specific variable
name.  If not otherwise specified, the message object will be
bound to the \verb|message| variable. 

\begin{codelisting}
<receive \dots>
    \hired{<python:correlator test="message.getPoId() == poId"/>}
</receive>
<receive \hired{bind="command"} \dots>
    \hired{<python:correlator test="command.getAction() == 'accept'"/>}
</receive>
\end{codelisting}

\begin{attrDefs}
test		&	required	&			& Java boolean correlation expression. \\
\end{attrDefs}

\section{\tag{python:action}}

The \tag{python:action} tag may contain a block of Python
to be interpreted as an atomic action.  The process case
attributes and other action attributes are bound to similarly
named variables within the block.

\begin{codelisting}
\hired{<python:action>
    \dots
</python:action>}
\end{codelisting}
